---
title: "**Introduction to Open Data Science, chapter6**"

subtitle: "**Analysis of Longitudinal data**"

author: "Yu Ren"
date: "12.8.2022"

output: 
  html_document:
    theme: flatly
    highlight: haddock
    toc: true
    toc_depth: 2
    number_section: false
---
Exercise 1
In this exercise, I will use the RATS dataset, and implement the analyses of Chapter 8 of MABS.
Prepare the packages
```{r}
library(dplyr)
library(tidyr)
```

read the RATS data set
```{r}
RATS <- read.table("data/ratsl.txt", sep = ",", header = T)

RATS$ID <- factor(RATS$ID)
RATS$Group <- factor(RATS$Group)

glimpse(RATS)
```
The dataset has 176 rows and 5 columns.

```{r}
# Plot the RATSL data
library(ggplot2)
ggplot(RATS, aes(x = Time, y = Weight, group = ID)) +
  geom_line(aes(linetype = Group))+
  scale_x_continuous(name = "Time(days)", breaks = seq(0, 60, 10))+
  scale_y_continuous(name = "Weight (grams)")+
  theme(legend.position = "top")
```
In the visualisation, it can be seen very visually that the weight of each rat increases slightly with time.
Next the weights were standardised and the search for the presence of this change continued.
```{r}
library(dplyr)
library(tidyr)
# Standardise the variable bprs
RATS_std <- RATS %>%
  group_by(ID) %>%
  mutate(stdweight = (Weight - mean(Weight))/sd(Weight)) %>%
  ungroup()

# Glimpse the data
glimpse(RATS_std)

# Plot again with the standardised bprs
library(ggplot2)
ggplot(RATS_std, aes(x = Time, y = stdweight, linetype = Group)) +
  geom_line() +
  scale_x_continuous(name = "Time (days)", breaks = seq(0, 64, 7)) +
  facet_grid(. ~ Group, labeller = label_both) +
  scale_y_continuous(name = "standardized weight") 
```
After standardizing the data, it is easy to see that the weighting increases gradually.

Then summary measure analysis of the data set
```{r}
# Number of subjects (per group):
n <- 11

library(dplyr)
library(tidyr)

# Summary data with mean and standard error of Weight by Group and Time 
RATSS <- RATS %>%
  group_by(Group, Time) %>%
  summarise( mean = mean(Weight), se = sd(Weight)/sqrt(n) ) %>%
  ungroup()

# Glimpse the data
glimpse(RATSS)

# Plot the mean profiles
library(ggplot2)
ggplot(RATSS, aes(x = Time, y = mean, col = Group)) +
  geom_line() +
  #geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=0.3) +
  theme(legend.position = "right") +
  scale_y_continuous(name = "mean(Weight) +/- se(Weight)") +
  scale_x_continuous(name = "Time (days)", breaks = seq(1, 64, 7))
```

Applying the summary measure approach
```{r}
library(dplyr)
library(tidyr)
# Summary data with mean and standard error of rats by treatment and week 
RATSS1 <- RATS %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()

# Glimpse the data
glimpse(RATSS)

# Draw a boxplot of the mean versus treatment
library(ggplot2)
ggplot(RATSS1, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(weight), days 1-64")

# Create a new data by filtering the outlier and adjust the ggplot code the draw the plot again with the new data
RATSS2 <- RATSS %>%
  filter(mean < 580)

library(ggplot2)
ggplot(RATSS2, aes(x = Group, y = mean))+
  geom_boxplot()+
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 4, fill = "white")+
  scale_y_continuous(name = "mean(weight), days 1-64")
```

Anova
```{r}
library(dplyr)
library(tidyr)
# Add the baseline from the original data as a new variable to the summary data
RATSS3 <- RATSS %>%
  mutate(baseline = RATS$day0)

# Fit the linear model with the mean as the response 
fit <- lm(mean ~ Group, data = RATSS3)

# Compute the analysis of variance table for the fitted model with anova()
anova(fit)
```

Exercise 2
In this exercise, I will use BPRS dataset.
The explain of BPRS from MABS4IODS:"Here 40 male subjects were randomly assigned to one of two treatment groups and each subject was rated on the brief psychiatric rating scale (BPRS) measured before treatment be- gan (week 0) and then at weekly intervals for eight weeks. The BPRS assesses the level of 18 symptom constructs such as hostility, suspiciousness, halluci- nations and grandiosity; each of these is rated from one (not present) to seven (extremely severe). The scale is used to evaluate patients suspected of having schizophrenia."

read the data set
```{r}
BPRS <- read.table("data/bprsl.txt", sep = ",",  header = T)
# Access the packages dplyr and tidyr
library(dplyr)
library(tidyr)

# Extract the week number
BPRS <-  BPRS %>% 
            mutate(week = as.integer(substr(BPRS$weeks, 5, 5)))

# Take a glimpse at the BPRSL data
glimpse(BPRS)
```

```{r}
# Plot the data
library(ggplot2)
ggplot(BPRS, aes(x = week, y = bprs, group = subject)) +
  geom_line() +
  theme(legend.position = "top") +
  facet_grid(. ~ treatment, labeller = label_both)
```

Holding on to independence: The Linear model
```{r}
# create a regression model BPRS_reg
BPRS_reg <- lm(bprs ~ week + treatment, data = BPRS)

# print out a summary of the model
summary(BPRS_reg)
```
The Random Intercept Model
```{r}
# access library lme4
library(lme4)

# Create a random intercept model
BPRS_ref <- lmer(bprs ~ week + treatment + (week | subject), data = BPRS, REML = FALSE)

# Print the summary of the model
summary(BPRS_ref)
```
Slippery slopes: Random Intercept and Random Slope Model
```{r}
# create a random intercept and random slope model
library(lme4)
BPRS_ref1 <- lmer(bprs ~ week + treatment + (week | subject), data = BPRS, REML = FALSE)

# print a summary of the model
summary(BPRS_ref1)

# perform an ANOVA test on the two models
anova(BPRS_ref1, BPRS_ref)
```
Time to interact: Random Intercept and Random Slope Model with interaction
```{r}
# create a random intercept and random slope model with the interaction
library(lme4)
BPRS_ref2 <- lmer(bprs ~ week + treatment + (week | subject) + week * treatment, data = BPRS, REML = FALSE)

# print a summary of the model
summary(BPRS_ref2)

# perform an ANOVA test on the two models
anova(BPRS_ref2, BPRS_ref1) 

# draw the plot of BPRS with the observed Weight values
library(ggplot2)
ggplot(BPRS, aes(x = week, y = bprs, group = subject)) +
  geom_line(aes()) +
  scale_x_continuous(name = "Time (weeks)") +
  scale_y_continuous(name = "Observed BPRS") +
  theme(legend.position = "top") 

# Create a vector of the fitted values
Fitted <- fitted(BPRS_ref2)

library(dplyr)
library(tidyr)
# Create a new column fitted to RATSL
BPRS <- BPRS %>%
  mutate(Fitted = Fitted)

# draw the plot of RATSL with the Fitted values of weight
library(ggplot2)
ggplot(BPRS, aes(x = week, y = Fitted, group = subject)) +
  geom_line(aes()) +
  scale_x_continuous(name = "Time (weeks)", breaks = seq(0, 60, 20)) +
  scale_y_continuous(name = "Fitted BPRS") +
  theme(legend.position = "top")

```









